---
title: "data_manipulation with 'dplyr"
output: github_document
date: "2022-09-27"
---

```{r}
library(tidyverse)

options(tibble.print_min = 3)

litters_data = 
  read_csv("./data/FAS_litters.CSV")
litters_data = 
  janitor::clean_names(litters_data)
#janitor function 可以帮助clean 文件里面的variable name里面包含的特殊字符，意思就是把variable name统一格式，整理数据！

pups_data = read_csv("./data/FAS_pups.csv")
pups_data = janitor::clean_names(pups_data)

```

###`select`

you can use `select` function to select columns when you have large datasets and specify the columns you want to keep by naming all ofthem. 

```{r}
select(litters_data, group, litter_number, gd0_weight, pups_born_alive)
```

you can also write the select in the form of below code, it's more succinct

```{r}
select(litters_data, group:pups_born_alive)
```

you can also specify columns you'd like to remove:

```{r}
select(litters_data, -pups_survive)
select(litters_data, -pups_survive, -group)
```

you can rename variables as part of this process: (new name is on the left side of  “=” )

```{r}
select(litters_data, GROUP = group, LiTeEr_NuMbEr = litter_number)
```

select 以gd为开头名字的variables
```{r}
select(litters_data, starts_with("gd"))

```

`everything()' is handy for re-organizing columns without discarding anything 

```{r}
select(litters_data, litter_number, pups_survive, everything())
```

you will often filter using comparison operators (`>`,`>=`,`<`,`<=`,`==`,and`!=`). You may also use `%in%` to detect if values appears in a set, and `is.na()`to find missing values. The results of comparisons are logical -- the statement is `TRUE` or `FALSE` depending on the values you compare -- and can be combined with other compasisons using the logical operators`&` and `|`, or negated using `!`. 

Some ways you might filter the litters data are: 

* `group %in% c("Con7", "Con8")`

```{r}
filter(litters_data, gd_of_birth == 20)
```

```{r}
filter(litters_data, group == "Con7" & gd_of_birth == 20)
```

* `drop_na(litters_data)` will remove any row with a missing value 
* `drop_na(litters_data, wt_increase)` will remove rows for which `wt_increase` is missing. 

### `mutate` 
we use `mutate` to create new variable 
```{r}
mutate(litters_data,
       wt_gain = gd18_weight - gd0_weight, 
       group = str_to_lower(group))

```


### arrange 

you can arrange the rows in your data according to the values in one or more columns: 

```{r}
head(arrange(litters_data, group, pups_born_alive), 10)
```

you can also sort in descending order if you'd like

```{r}
head(arrange(litters_data, desc(group), pups_born_alive), 10)
```

```{r}
litter_data = 
  read_csv(".data/FAS_litters.csv", col_types = "ccddiiii") %>%
  janitor::clean_names() %>%
  select(-pups_survive) %>%
  mutate(
    wt_gain = gd18_weight - gd0_weight,
    group = str_to_lower(group)) %>%
  drop_na(wt_gain)

litters_data
```


